# Week3



```{r include=FALSE}
alc <- read.table("data/alc.txt")
#alc1 <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep=",", header=TRUE)
library(dplyr)
keep_columns <- c("sex","age","famrel", "absences", "high_use")

# select the 'keep_columns' to create a new dataset
pp <- select(alc, one_of(keep_columns))
```

This week dataset combines two dataset that approach student achievement in secondary education of two Portuguese schools. Datasets provides info about students performance in two different subject. Both of these datasets includes same information regarding students backround. 


Names of the variables of the this week dataset:
```{r echo=FALSE}
colnames(alc)
```

I chose to study variables age, sex, absences and family relationship (famrel) more closely and their relationship with alcohol consumption. My hypothesis is that age, sex and absences correlates positively with alcohol consumption and famrel correlates negatively. In addition I expect that male students use more alcohol than female.



The distributions of the chosen variables:
```{r echo=FALSE}
library(tidyr)
# draw a bar plot of each variable
gather(pp) %>% ggplot(aes(value)) + geom_bar() + facet_wrap("key", scales = "free")
```

There is no normal distribution among my variables. The absences variable does not have any reosonable pattern. A typical student age is between 15-18 and he has good quality of family relationship. Gender is roughly equally distributed among students.

```{r echo=FALSE}
#keep_columns <- c("sex","age","famrel", "absences", "high_use")

# select the 'keep_columns' to create a new dataset
#pp <- select(alc, one_of(keep_columns))

# initialise a plot of high_use and age
g1 <- ggplot(alc, aes(x = high_use, y = age))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("age") + ggtitle("Student age by alcohol consumption")
```

It seems to be so that older students consume more alcohol than younger as I expected.

```{r echo=FALSE}
# initialise a plot of high_use and absences
g2 <- ggplot(alc, aes(x = high_use, y = absences))

# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("absences") + ggtitle("Student absences by alcohol consumption")
```

Students who have more absences seems to comsume more alcohol than students who do not skip classes, as I expected.


```{r echo=FALSE}
# initialise a plot of high_use and quality of family relationships
g2 <- ggplot(alc, aes(x = high_use, y = famrel))

# define the plot as a boxplot and draw it
g2 + geom_boxplot() + ylab("famrel") + ggtitle("Student quality of family relationships by alcohol consumption")
```

There is less high alcohol use among those students who have a good quality of family relationships.

```{r echo=FALSE}
# initialize a plot of 'high_use'
g2 <- ggplot(data = alc, aes(x = high_use, fill=sex))

# draw a bar plot of high_use by sex
g2 + geom_bar()
#g2 + facet_wrap("sex")
```

Thre are more male students among high alcohol consumers according to this graph.

All in all, these plots support my earlier hypothesis.


Summary of the fitted model:
```{r echo=FALSE}
# find the model with glm()
m <- glm(high_use ~ sex+ age + famrel + absences, data = alc, family = "binomial")

# print out a summary of the model
summary(m)

```

All the variables are statistically significant. Although the age variable is statistically significant only in 0.1 significance level. SexM and absences variables are statistically the most significant variables. Gender and quality of family relationships seems to have quite large impact on students high use of alcohol.

The coefficients of the model as odds ratios and confidence intervals for them:
```{r echo=FALSE}
# print out the coefficients of the model
# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI <- confint(m) %>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```

These results are in line with my earlier stated hypothesis.


```{r echo=FALSE}
# produce summary statistics by group
#alc %>% group_by(age, high_use) %>% summarise(count = n())
```

```{r include=FALSE}
# initialize plot with data and aesthetic mapping
#p1 <- ggplot(alc, aes(x = alc_use, y = famrel))

# define the visualization type (points)# add a regression line
#p2 <- p1 + geom_point()+ geom_smooth(method = "lm")


# add a main title and draw the plot

#p2
#my_model <- lm(high_use ~ famrel, alc)
#plot(my_model, which = c(1,2,5))
```


```{r echo=FALSE}

```

