# Week 2

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using 

```{r include=FALSE}
learn14 <- read.table("data/learning2014.txt")
#read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
```

This dataset consist answers to a survey which was done in course Introduction to Social Statistics, fall 2014. The survey questions were related to students learning approaches. Questions were divided to three different categories: deep learning, surface learning and strategic learning. In this dataset answers to those question categories has been compined to columns: deep, surf and stra. The dataset includes also students ages, gender, attitude (a sum of 10 questions related to students attitude towards statistics) and test scores. 

Structure and dimension of dataset
```{r echo=FALSE}
#Structure and dimension of dataset
str(learn14)
dim(learn14)
```


```{r echo=FALSE}
library(ggplot2)

# initialize plot with data and aesthetic mapping
p1 <- ggplot(learn14, aes(x = Attitude, y = Points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# add a regression line
p3 <- p2 + geom_smooth(method = "lm")

# add a main title and draw the plot
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4
```


