# Week 2

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using 

```{r include=FALSE}
learn14 <- read.table("data/learning2014.txt")
#read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
```

This dataset consist answers to a survey which was done in course Introduction to Social Statistics, fall 2014. The survey questions were related to students learning approaches. Questions were divided to three different categories: deep learning, surface learning and strategic learning. In this dataset answers to those question categories has been compined to columns: deep, surf and stra. The dataset includes also students ages, gender, attitude (a sum of 10 questions related to students attitude towards statistics) and test scores. 

Structure and dimension of the dataset:
```{r echo=FALSE}
#Structure and dimension of data
str(learn14)
dim(learn14)
```


```{r eval=FALSE, include=FALSE}
# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

# initialize plot with data and aesthetic mapping
p1 <- ggplot(learn14, aes(x = Attitude, y = Points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# add a regression line
p3 <- p2 + geom_smooth(method = "lm")

# add a main title and draw the plot
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4
```

Graphical overview of the data and summaries of the variables in the data:

```{r echo=FALSE}
# draw a scatter plot matrix of the variables in learning2014.
# [-1] excludes the first column (gender)
#pairs(learn14[-1])
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(learn14, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p
```

The graphical overview shows us that variables attitude, stra and surf have normal distribution. Variable age disturibution is clearly toward left. Distributions of variables deep and points are toward right, allthough points distribution have also a quite fat tail on left side.

Attitude variable is correlated with points in both genders and the correlation is about 0.43. Also variable age is correlated with points but only among males. Variable surf has some correlation with variables attitude, deep and stra among males. All other variables have only little correlation or no correlation at all.

I chose variables attitude, stra and surf as explanatory variables and fitted a regression model where exam points is the target variable. 

Summary of the fitted model: 

```{r echo=FALSE}
my_model <- lm(Points ~ Attitude + stra + surf, learn14)
summary(my_model)
```
Explain and interpret the statistical test related to the model parameters.

The statistical test and p-value measures that how likely the estimate is zero. Eg. if the p-value is very low then most likely the estimate is not zero and we can say that it is statistically significant.

In my model only attitude is statistically significant so I run regresion again using only that explanatory variable.

Summary of the new model:
```{r echo=FALSE}
my_model <- lm(Points ~ Attitude, learn14)
summary(my_model)
```

4. Using a summary of your fitted model, explain the relationship between the chosen explanatory variables and the target variable (interpret the model parameters). Explain and interpret the multiple R squared of the model.

The estimate of the attitude variable is ~0.35 and it is statistically significant. This means that points and attitude are in relationship, where one additional attitude point gives 0.35 points more in exam according this model.

The multiple R-squered of this model is 0.19, which means that this model explain about 20% of changes in dependent variable (exam points). In other words, according this modele attitude towards statistics explains 20 prosent of differencies in Introduction to Social Statistics exam results.


```{r echo=FALSE}
par(mfrow = c(2,2))
plot(my_model, which = c(1,2,5))
```

5. Produce the following diagnostic plots: Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage. Explain the assumptions of the model and interpret the validity of those assumptions based on the diagnostic plots. 


