# Week 5 - Dimensionality reduction techniques



```{r include=FALSE}
human_ <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep  =",", header = T)
library(tidyr)
library(ggplot2)
library(corrplot)
library(dplyr)
library(GGally)
library(FactoMineR)
```
### Description of the dataset
#### Summaries of the variables in the data

```{r echo=FALSE}
summary(human_)
```

```{r echo=FALSE}
# draw a bar plot of each variable
gather(human_) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_histogram() + ggtitle("Distributions of variables")
```
Distributions of variables Ado.birrth, GNI, Parli.F and Mat.Mor are clearly skewed towards left. Edu.Exp has quite normal distribution. Variables Life.Exp and Labo.FM have the distuributions that are skewed towards right.

#### Correlations between variables

```{r echo=FALSE}
# calculate the correlation matrix and round it
cor_matrix<-cor(human_) %>% round(2)
# visualize the correlation matrix
corrplot(cor_matrix, order = "hclust", addrect = 2, type = "upper",tl.col = "black", tl.srt = 45, tl.cex = 0.7)
```
Variables rad and tax have the highest correlation (0,91). Other variable pairs that have high correlation (in abosolute value greater than 0,7) are "indus,nox", "indus,tax", "nox,age", "rm,mediv", "indus,dis", "nox,dis", "age,dis" and "lstat,medv".



### Principal component analysis (PCA) on the not standardized data

```{r echo=FALSE}
#2
# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_)

# create a summary of pca_human
s <- summary(pca_human)

# rounded percetanges of variance captured by each PC
pca_pr <- round(100*s$importance[2, ], digits = 1)

# print out the percentages of variance
pca_pr

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex=c(0.8, 1))
```

### PCA on the standardized data

```{r echo=FALSE}
#3
# standardize the variables
human_std <- scale(human_)

# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human_std)

# create a summary of pca_human
s <- summary(pca_human)

# rounded percetanges of variance captured by each PC
pca_pr <- round(100*s$importance[2, ], digits = 1)

# print out the percentages of variance
pca_pr

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex=c(0.8, 1))
```



### Tea time !
#### visualize the dataset
```{r echo=FALSE}
# 5
# load the data
data("tea")
# the tea dataset and packages FactoMineR, ggplot2, dplyr and tidyr are available

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# look at the summaries and structure of the data
summary(tea_time)
str(tea_time)
dim(tea_time)

# visualize the dataset
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) + geom_bar()

```
#### Summary of the model
```{r echo=FALSE}
# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
```
